Waving and swaying

### 2014-02-21 Paper/grass.tex
updatedFeb21

\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{slashbox}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{dashrule}
\usepackage{amsmath}

\usepackage{dcolumn}% Align table columns on decimal point

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{ifthen}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}
\graphicspath{ {images/} }

@@

The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable either through a mechanism of Kelvin-Helmholtz instability associated with 
inflection point or via shear instability of flow above the grass.

\end{abstract}
\maketitle

@@

Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.

@@

Due to these considerable differences between aquatic and marine flow the analysis in this paper is applicable only to marine canopies  

@@

different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to Kelvin-Helmholtz
instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.

@@

several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self consistent 
theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical Reynolds number (Grizzle 1996) 
for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires further investigation for better understanding of
phenomenon.

@@

In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grasses 
shows that existence of strong shear near canopy top does not lead to Kelvin Helmholtz instability. We also show instead that mo-nami is caused by instability 
of fluid flow above vegetation coupled with the flow through canopy. We further show that inclusion of drag due to grass predicts a threshold condition for waving where threshold is
characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.  

@@

While Drag exerted by canopy is believed to play 
a significant role in the development of coherent vortices leading to monami, flexibility of grass provide striking visualization of these structures without significantly affecting the
dynamics of flow (Raupach 1996, Ghisalberti 2002). For this reason we replace grass with stiff structure to investigate dynamics leading to generation of 
coherent vortices. We use a mean field model for the coupling between the flow and the canopy. 
%The drag force per unit volume experienced by fluid flow is believed to be dependent on
%mean frontal area of grass. We approximate mean frontal area of grass field of height $h$ with $dN_g/m^2$. Where $d$ is average width of grass and $N_g$ the grass number density per unit area 
Drag force exerted by the grass is approximated via a continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as
%We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
%taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
%shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
% and solving force balance and momentum balance for grass  

@@

@@

normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. We further expect $C_N$ to
not change along the height of the grass and take $C_N$ to be constant. 

@@

A solution of equation(1) gives uniform steady flow which we expect to be unstable leading to generate coherent eddies. To understand the instability characteristic associated with this
steady flow termed as base flow we perform linear stability analysis.  

@@

This base state flow is obtained by solving steady state form of constant pressure gradient driven flow of equation(1)
$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface Fig(1). Flow consist of a region of approximately constant velocity
$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ arising from balance of drag force with pressure gradient with in grass and a parabolic velocity profile in unvegetated region due to 
balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top give rise to two boundary layer in the flow profile one 
near bottom and near the tip of the grass which is conventionally known as shear layer.
%eads to a 
%constant velocity $U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$ while balance of viscous force with pressure gradient above grass provide a parabolic velocity profile.
%The base flow profile consist of two boundary layer one near the bottom of grass due to no slip condition and other near tip of canopy known as shear layer due to vertical discontinuity 
%of drag force. Width of boundary layer near bottom of grass can be obtained by balance of viscous force with drag and pressure gradient forces which gives boundary layer
%width $\sim \sqrt{\frac{\mu}{\rho C_NdN_g U_g}}$. 
Thickness of shear layer near top of grass is obtained by balance of viscous force 

@@

width $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$ where $U_0 = \frac{dP/dxH^2}{\mu}$ is scale for velocity.

\begin{figure}[htb!]

@@

@@

%\scriptstyle{

@@

 %u_t+Uu_x+vU_y &= \frac{1}{\rho}p_x+\frac{\mu}{\rho}\left(u_xx+u_yy \right) -2C_NdN_gUu\\

@@

%}

@@

@@

@@

non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 

@@

&-\frac{\del}{\del y}\left(\frac{2D_{drag}}{R_e}U\phi_y\right)

@@

Variation of critical wave number with shear layer thickness for different grass height is shown in Fig-3.
% shows 
%this critical number for different grass density and grass height along with experimental threshold observed by Grizzle (1996)
\begin{figure}[htb!]

@@

The following features can be noted from Fig(3). For the certain grass height region of linear variation of wavenumber with inverse shear layer thickness suggest that the flow is subject 
to Kelvin-Helmholtz instability above a threshold shear thickness. Region of constant wave number suggest that length scale of the instability is no longer governed by shear layer
thickness in this this regime, rather it is governed by flow above the grass, which can be understood by asymptotic analysis in regime of small shear layer thickness and large Reynolds 
number
\begin{figure}[htb!]

@@

@@

%\newline
Collapse of $\phi$ for different shear thickness in the regime of constant critical wavenumber $k$ suggest that we have obtained the asymptotic solution in limit of small shear layer 
thickness. Noting that non-dimensional flow velocity $U \sim \delta^{3/2}$ inside the grass and $R_{e} \gg 1$, we can simplify the equation-(4) considerably both with in the grass and 
above the grass canopy. Inside the grass equation simplifies into $\sigma\left(D^2-k^2\right)\phi = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$. Together with simplified
force balance inside the grass canopy and constant value of $k\approx O(1)$, we can estimate that critical Reynolds number $R_e \sim \delta^{-3/2}$ for small shear layer thickness Fig(2) 
\begin{figure}[htb!]

@@

@@

\begin{figure}[htb!]
\includegraphics[scale=0.34]{Observed_vs_calculated}
\caption{Comparison between experimently oberved frequency and predicted frequency}
\end{figure}

>>>>

%\documentclass[aps,prl,twocolumn,showpacs,superscriptaddress,groupedaddress,10pt]{revtex4-1}  % for review and submission
\documentclass[aps,twocolumn,floatfix,prl,10pt]{revtex4-1}
%\documentclass[letterpaper,10pt,prl,twocolumn,aps]{revtex4-1}
%\documentclass[aps,preprint,floatfix,prl]{revtex4-1}
%\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{slashbox}
\usepackage{color}
\usepackage{longtable}
\usepackage{array}
\usepackage{dashrule}
\usepackage{dcolumn}% Align table columns on decimal point
%\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{ifthen}
\usepackage{amsthm} % Theorem Formatting
\usepackage{amssymb}	% Math symbols such as \mathbb
\usepackage{calrsfs}
\usepackage{subfigure}
\graphicspath{ {images/} }

@@

The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed.
\end{abstract}
\maketitle

@@

Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami (Inoue 1955, Raupach 1996) have also been observed.

@@

Due to these considerable differences between aquatic and marine flow, our analysis of flow in presence of flexible grass in a bounded domian is applicable only to marine canopies. 

@@

different amounts of drag experienced by fluid in and above the canopy. This shear layer is assumed to become unstable to coherent vortices through a mechanism similar to 
Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses is manifested in their large amplitude synchronous oscillations.

@@

several aspect of existing theory remain unexplained. First, linear stability analysis of base flow does not take into account influence of drag due to grass hence not a self 
consistent theory. Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical 
Reynolds number (Grizzle 1996) for monami is much higher $\approx O(1000)$. These drawbacks of existing theory suggest that flow through vegetation requires 
further investigation for better understanding of phenomenon.

@@

In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grass
shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz 
instability associated with strong shear near the canopy top but also through shear instability of flow above grass canopies arising due to existence of flow boundary.
We further show that inclusion of drag due to grass predicts a threshold condition for waving where 
threshold is characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.

@@

The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depends weakly on flexibility of grass
unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as

@@

\label{navier-stokes}

@@

normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field such as marine
both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N \& N_g$ to be constants.

@@

The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. steady state flow driven by   

@@

constant pressure gradient satisfies
$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. A sample solution is shown in figure (1) and 
compared with mean flow observed in lab scale experiment. Flow consist of a region of approximately constant velocity
\small$U_g \sim \sqrt{\frac{dP/dx}{\rho C_N dN_g}}$\normalsize arising from balance of drag force with pressure gradient with in grass and a parabolic velocity 
profile in unvegetated region due to balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top 
give rise to two boundary layer in the flow profile one near bottom and other near the tip of the grass. The boundary layer near the canopy top is identified 
as shear layer in previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
In our model thickness of shear layer near top of grass is obtained by balance of viscous force 

@@

width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity.

@@

\small

@@

@@

@@

@@

\normalsize

@@

\begin{figure}[htb]

@@

non-dimensional drag. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, we get generalized orr-sommerfield equation 

@@

&-\frac{\del}{\del y}\left(\frac{2}{R_e}\frac{H^3}{\delta^3}U\phi_y\right)
\label{Orr-somerfield}

@@

Variation of critical wave number with shear layer thickness for different grass height is shown in Figure \ref{K_vs_shear}
\begin{figure}[htb]

@@

Variation of critical wavenumber with shear layer thickness exhibits a region of linear variation of critical wavelength with shear layer thickness showing flow instability arising from
a mechanism similar to Kelvin-Helmholtz as hypothesized earlier (Ghisalberti, Raupach). However the region of almost no variation of wavelength for a given submergence ratio with 
shear layer thickness shows the existence of another regime where flow instability does not arise from Kelvin-Helmholtz instability of shear layer rather flow instability is dominated by
the instability of flow in unvegetated region. The existence of a regime dominated by shear instability of flow above grass is further corroborated by the collapse of eigen-modes 
for different shear layer thickness into a single profile as shown in Figure \ref{Asymptotic_mode}. The regime of shear instability of flow above the grass further predicts the
asymptotic variation of critical Reynolds number $R_e \sim \delta^{-3/2}$ with shear layer thickness which can be understood by asymptotic analysis of \eqref{Orr-somerfield}.
\begin{figure}[htb]

@@

\label{K_vs_shear}

@@

Noting that for large drag, non-dimensional flow velocity $U \sim (\frac{\delta}{H})^{3/2}$ inside the grass and $R_{e} \gg 1$, equation \eqref{Orr-somerfield}
can be simplified considerably both with in the grass and above the grass canopy. As $U_{yy}\sim 0$ inside the grass and $\frac{1}{R_e} \ll 1$ for the regime of unvegetated flow instability
 the equation within the grass simplifies into
into $\sigma\left(\phi_{yy}-k^2\phi\right) = -\frac{\del}{\del y}\left(\frac{2\phi_y}{\delta^{3/2}R_e}\right)$ while in the unvegetated region it simplifies into Rayleigh equation 
$ \left(\sigma+ikU\right) \left(D^2-k^2\right)\phi =  ikU_{yy}\phi$.
\begin{figure}[htb]

@@

\label{Asymptotic_mode}

@@

\begin{figure}[htb]
\includegraphics[scale=0.34]{Observed_vs_calculated}
\caption{Comparison between experimently oberved frequency and predicted frequency}
\begin{figure}[htb]
\subfigure{\includegraphics[]{Vivoni_Fig3_6_zero_shear_match}}
\subfigure{\includegraphics[]{Vivoni_Fig3_7_zero_shear_match}}
\caption{Comparison between observed and calculated velocity profile}
%\begin{figure}[htb]
%\includegraphics[]{Vivoni_Fig3_7_hb_14match}
%\caption{Comparison between observed and calculated velocity profile,hb=14}
%\end{figure}
%\begin{figure}[htb]
%\includegraphics[]{Vivoni_Fig3_7_hb_17match}
%\caption{Comparison between observed and calculated velocity profile,hb=17}
%\end{figure}
